---
title: "Jiaming Ren 217218863 Assignment 1"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load libraries}
library(tidyverse)
library(ggplot2)
library(stringr)
library(tidyr)
library(data.table)
library(dplyr)
library(naniar)
```
# Q1
## read the World_development_Indicators.csv
```{r read dataset}
ds_worldDI = read_csv("World_Development_Indicators.csv")
```
## display head of dataset
```{r }
##display first 5 records
head(ds_worldDI)
```

## columns and rows of dataset
```{r rows, columns and structure of dataset}
#number of columns
print("number of columns")
ncol(ds_worldDI)

#number of rows
print("number of rows")
nrow(ds_worldDI)

#the structure of dataset
print("the structure of dataset")
str(ds_worldDI)
```
# Q2 
## 1.Select unique value of series name and code
## 2.count unique series 
## 3.print first 9 rows
## 4.find how many contain the keyword fertility
```{r }
#Total unique values for the Columns Series Name and code
unique_series <- unique(na.omit(ds_worldDI[c("Series Name","Series Code")]))

unique_series

#get the length of unique series
len_unique_series <- unique_series[1] %>% count()

len_unique_series

```

```{r }
#print the first 9 unique series
unique_series %>% head(9)
```

```{r }
#get series that contain the keyword Fertility
fertility <- ds_worldDI[str_detect(ds_worldDI$"Series Name", regex("fertility",ignore_case = TRUE)), ] 
fertility %>% length()

#get unique series that contain the keyword Fertility
unique_fertility <- unique_series %>% filter(`Series Name` == regex("Fertility", ignore_case = T))
#row numbers of rows that contain fertility
unique_fertility %>% length()

```

## (1) top 5 countries
### 1. get all rows for GDP (current US$)
### 2. sort the dataset and get top 5 countries with highest GDP (current US$) in 2019
## (2) bottom 5 countries
### 3. sort the dataset and get bottom 5 countries with hightest GDP (current US$) in 2019
### 4. only keep the GDP data related to the 10 countries

```{r }
#select country name and 2019.00 columns
gdp_2019 <- ds_worldDI %>% filter(ds_worldDI$"Series Name" == "GDP (current US$)") %>% select("Country Name", "2019.00")

#sort the rows by the column 2019.00 and select bottom 5 countries
bot <- gdp_2019 %>% arrange(gdp_2019$"2019.00") %>% head(5)
bot
#sort the rows by the column 2019.00 and select top 5 countries
top <- gdp_2019 %>% arrange(desc(gdp_2019$"2019.00")) %>% head(5)
top
#combined two top and bot datasets
filtered_countries <- rbind(top,bot)$"Country Name"

#filter and keep the GDP data related to the 10 coutries
filtered_ds_worldDI <- ds_worldDI %>% filter(ds_worldDI$"Country Name" %in% filtered_countries)
filtered_ds_worldDI
```

# Q4
## 1. count how many columns contain missing values
## 2. list each of these columns with the corresponding missingess percentages
```{r }
#visualize each of year columns with the corresponding missing percentages  
vis_miss(filtered_ds_worldDI[,c(-1,-2,-3,-4)])

```

```{r }
#convert the table from wide to long. group by year and value
year_value <- filtered_ds_worldDI %>% gather(Year, Value, -"Country Name", -"Country Code", -"Series Name",  -"Series Code")

#print into a table about how much missing value for each year in percentage
year_value %>% 
  group_by(Year) %>% 
  summarise(Percentage = mean(is.na(Value)*100), Na_rows= is.na(Value) %>% sum(), total_rows = Value %>% length())
```
#Q5
## 1. display statistical infromation. min max and mean of the gdp from 2010 - 2019
```{r }
#statistical data only for min and max and mean
year_value %>% 
  group_by(Year) %>% 
  summarise(Min =Value %>% min(na.rm = T), Max = Value %>% max(na.rm=T), Mean = Value %>%  mean(na.rm=T)) %>% 
  filter(Year == c("2010.00","2011.00","2012.00","2013.00","2014.00","2015.00","2016.00","2017.00","2018.00","2019.00"))

#all statistical from 2010 - 2019
summary(filtered_ds_worldDI[55:length(colnames(filtered_ds_worldDI))])
```
# Q6
## 1.get all china value from the year_value table
## 2.print max gdp for china 
## 3.print min gdp for china
```{r }
#select china and gdp rows
China_df<-year_value %>% 
  filter(`Country Name` == "China") %>% 
  filter(`Series Name` == "GDP (current US$)")

print("max gdp:")
China_df$Value %>% max()
print("min gdp:")
China_df$Value %>% min()
  
```
# Q7
## 1.filter all GDP information for 10 selected countries (top5 and bot5).
## 2.plot GDP for 10 selected countries over all years.
## 3.answer question what is one advantage and one idsadvantage of including all these countries in one plot
```{r }
#filter to get GDP (current US$) series for each Country in each year
ten_gdp <- year_value %>% 
  filter(`Series Name` == "GDP (current US$)")

#plot all ten countries in one graph
ten_gdp %>% 
  ggplot(aes(Year, Value, group=`Country Name`, color = `Country Name`))+
  geom_line() +
  geom_point()

#the advantage is that it is easier to compare each country in one graph
#the disadvantage is that the top 5 and bot 5 countries has dramastic different of GDP. From the graph, you can see the bot countries are stacked into one line.
```
# Q8
## 1.create a new column and name it as Period
## 2.assign Period factors 
## 3.create a table that shows mean of each period for each country
## 4.plot graph for each country
```{r }
#add a Period columns
ten_gdp$Period<-''

#label Periods
ten_gdp$Period[ten_gdp$Year <1970 &ten_gdp$Year >=1960]<- "Period1"
ten_gdp$Period[ten_gdp$Year <1980 &ten_gdp$Year >=1970]<- "Period2"
ten_gdp$Period[ten_gdp$Year <1990 &ten_gdp$Year >=1980]<- "Period3"
ten_gdp$Period[ten_gdp$Year <2000 &ten_gdp$Year >=1990]<- "Period4"
ten_gdp$Period[ten_gdp$Year <2010 &ten_gdp$Year >=2000]<- "Period5"
ten_gdp$Period[ten_gdp$Year <2020 &ten_gdp$Year >=2010]<- "Period6"

#create a table that shows mean of each period for each country
average_period <- ten_gdp %>% group_by(`Country Name`, Period) %>% summarise(average = Value %>% mean(na.rm = T), .groups = 'drop')
average_period

#plot graph for each country
average_period %>% 
  ggplot(aes(Period, average,color=`Country Name`,group=`Country Name`))+
  geom_line()+
  facet_wrap(~`Country Name`,scale="free")

```

# Q9
## 1.subset America with GDP
## 2.swap wide table to long table
## 3.assign mean value to all NA
```{r }
#subset America with GDP
America <- filtered_ds_worldDI %>% filter(`Country Name` == "United States") %>% filter(`Series Name` == "GDP (current US$)")

#swap wide table to long table
America <- America %>% gather(Year,Value, -"Country Name", -"Country Code", -"Series Name", -"Series Code")

#assign mean value to NA
America$Value[is.na(America$Value)] <- America$Value %>% mean(na.rm=T)

America
```